### 2014-06-03 Paper.tex
Adding paper.*

\documentclass[10pt]{article}
\usepackage{url,graphicx,tabularx,array}%,geometry} %,fullpage,amsfonts,amsmath}
\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd,mathabx,ulem}
\usepackage{fullpage}
\usepackage{lastpage}
\usepackage{enumerate}
\usepackage{fancyhdr}
\usepackage{mathrsfs}
\usepackage[margin=1.5cm]{geometry}
\usepackage{listings}
\usepackage{mathabx}
\usepackage{tikz,pgfplots}
\usepackage{enumitem}
\usepackage{graphicx,wrapfig}
\usepackage{color,hyperref,endnotes}

\usetikzlibrary{arrows,shapes}
\usepackage{tikz-qtree}

%\setlength{\parskip}{1ex} %--skip lines between paragraphs
\setlength{\parindent}{0pt} %--don't indent paragraphs
\setlist{nolistsep}
%-- Commands for header

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}
\definecolor{lightgray}{rgb}{.9,.9,.9}
\definecolor{darkgray}{rgb}{.4,.4,.4}
\definecolor{purple}{rgb}{0.65, 0.12, 0.82}

% \renewcommand{\title}[1]{\textbf{#1}\\}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\gray}[1]{\textcolor{gray}{#1}}

\lstset{basicstyle=\footnotesize\ttfamily, 
	tabsize=2,
	language=Python,
	backgroundcolor=\color{lightgray},
	commentstyle=\color{mygreen},
	keywordstyle=\color{blue},
	stringstyle=\color{red}\ttfamily,
	showstringspaces=false,
	breaklines=true
}

\title{CS246: Twitter Purification}
\author{Wen Shi, Zijun Xue, Jennifer Zhang}
\date{June 2014}
\begin{document}
\maketitle

\section*{Abstract}
\section*{Introduction}
\subsection*{Motivation}
\subsection*{Overview}
\section*{Implementation}
\subsection*{Dictionary}
TODO: Jennifer
\subsection*{Single word correction}
\subsubsection*{Abbreviations: single word}
TODO: Jennifer
\subsubsection*{Squeeze}
TODO: Zijun
\subsubsection*{Edit distance}
TODO: Zijun
\subsubsection*{Phonetic candidates: Soundex}
TODO: Shi Wen
\subsubsection*{Phonetic candidates: Metaphone}
TODO: Shi Wen
\subsubsection*{Letter similarity: Viterbi}
TODO: Jennifer
\subsubsection*{Word frequency scaling}
TODO: Jennifer
\subsubsection*{Abbreviations: phrase}
TODO: Jennifer
\subsection*{Bigram correction}
TODO: add some subsections here?
\section*{Discussion and Evaluation}
\subsection*{Single word results}
\subsection*{Bigram results}
\section*{Future work}
TODO: listing any more aspects where it didn't work well
\subsubsection*{Unconventional word tokenization}
\subsubsection*{Obscure vocabulary or slang}
\subsubsection*{Proper nouns and acronyms}
\subsubsection*{Abbreviations}
\subsubsection*{Dictionary}

\end{document}